Reanalysis Step 8: Patient-Level Table 1

### 2025-09-05 reanalysis-8-table1.qmd
table1

  # Add total row
  n_lapse <- as.numeric(gsub(" .*", "", table_data$totals$Lapse))
  n_no_lapse <- as.numeric(gsub(" .*", "", table_data$totals$NoLapse))
  n_total <- as.numeric(gsub(" .*", "", table_data$totals$Total))

  export_df <- rbind(export_df, data.frame(
    variable = "Total",
    lapse_value = n_lapse,
    lapse_prop = as.numeric(gsub(".*\\((.*)%.*", "\\1", table_data$totals$Lapse)),
    lapse_sd = NA,
    no_lapse_value = n_no_lapse,
    no_lapse_prop = as.numeric(gsub(".*\\((.*)%.*", "\\1", table_data$totals$NoLapse)),
    no_lapse_sd = NA,
    total_value = n_total,
    total_prop = 100,
    total_sd = NA
  ))

  # Process categorical variables from table_data
  # This is a simplified version - you'll need to expand this based on your exact structure

  # Add treatments
  for (i in 1:3) {  # Lapse, No Lapse, Total
    if (i <= nrow(table_data$treatments$prp)) {
      # Extract values for each treatment
      # This needs more sophisticated parsing based on actual data structure
    }
  }

  # Add continuous variables
  for (var_name in c("baseline_VA_logMAR", "DCSI", "CCI")) {
    var_label <- switch(var_name,
      "baseline_VA_logMAR" = "Baseline VA (logMAR)",
      "DCSI" = "Diabetes Complication Severity Index",
      "CCI" = "Charlson Comorbidity Index"
    )

    # Parse mean ± SD format
    if (var_name == "baseline_VA_logMAR") {
      stats <- table_data$continuous$va$stat
    } else if (var_name == "DCSI") {
      stats <- table_data$continuous$dcsi$stat
    } else {
      stats <- table_data$continuous$cci$stat
    }

    # Parse "mean ± sd" format
    parse_mean_sd <- function(x) {
      parts <- strsplit(x, " ± ")[[1]]
      return(list(mean = as.numeric(parts[1]), sd = as.numeric(parts[2])))
    }

    lapse_stats <- parse_mean_sd(stats[1])
    no_lapse_stats <- parse_mean_sd(stats[2])
    total_stats <- parse_mean_sd(stats[3])

    export_df <- rbind(export_df, data.frame(
      variable = var_label,
      lapse_value = lapse_stats$mean,
      lapse_prop = NA,
      lapse_sd = lapse_stats$sd,
      no_lapse_value = no_lapse_stats$mean,
      no_lapse_prop = NA,
      no_lapse_sd = no_lapse_stats$sd,
      total_value = total_stats$mean,
      total_prop = NA,
      total_sd = total_stats$sd
    ))
  }

>>>>

  # Helper function to parse "mean ± sd" format
  parse_mean_sd <- function(x) {
    if (is.na(x) || x == "") return(list(mean = NA, sd = NA))
    parts <- strsplit(x, " ± ")[[1]]
    return(list(mean = as.numeric(parts[1]), sd = as.numeric(parts[2])))
  }

  # Helper function to parse "n (pct%)" format
  parse_n_pct <- function(x) {
    if (is.na(x) || x == "") return(list(n = NA, pct = NA))
    n <- as.numeric(gsub(" .*", "", x))
    pct <- as.numeric(gsub(".*\\((.*)%.*", "\\1", x))
    return(list(n = n, pct = pct))
  }

  # Add total row
  lapse_total <- parse_n_pct(table_data$totals$Lapse)
  no_lapse_total <- parse_n_pct(table_data$totals$NoLapse)
  total_total <- parse_n_pct(table_data$totals$Total)

  export_df <- rbind(export_df, data.frame(
    variable = "Total",
    lapse_value = lapse_total$n,
    lapse_prop = lapse_total$pct,
    lapse_sd = NA,
    no_lapse_value = no_lapse_total$n,
    no_lapse_prop = no_lapse_total$pct,
    no_lapse_sd = NA,
    total_value = total_total$n,
    total_prop = 100,
    total_sd = NA
  ))

  # Process categorical variables
  # Sex
  if (!is.null(table_data$sex)) {
    for (i in 1:nrow(table_data$sex)) {
      level <- table_data$sex$level[i]
      lapse_val <- parse_n_pct(table_data$sex$stat_lapse[i])
      no_lapse_val <- parse_n_pct(table_data$sex$stat_no_lapse[i])
      total_val <- parse_n_pct(table_data$sex$stat_total[i])

      export_df <- rbind(export_df, data.frame(
        variable = paste0("Sex_", level),
        lapse_value = lapse_val$n,
        lapse_prop = lapse_val$pct,
        lapse_sd = NA,
        no_lapse_value = no_lapse_val$n,
        no_lapse_prop = no_lapse_val$pct,
        no_lapse_sd = NA,
        total_value = total_val$n,
        total_prop = total_val$pct,
        total_sd = NA
      ))
    }
  }

  # Age groups
  if (!is.null(table_data$age)) {
    for (i in 1:nrow(table_data$age)) {
      level <- table_data$age$level[i]
      lapse_val <- parse_n_pct(table_data$age$stat_lapse[i])
      no_lapse_val <- parse_n_pct(table_data$age$stat_no_lapse[i])
      total_val <- parse_n_pct(table_data$age$stat_total[i])

      export_df <- rbind(export_df, data.frame(
        variable = paste0("Age_", level),
        lapse_value = lapse_val$n,
        lapse_prop = lapse_val$pct,
        lapse_sd = NA,
        no_lapse_value = no_lapse_val$n,
        no_lapse_prop = no_lapse_val$pct,
        no_lapse_sd = NA,
        total_value = total_val$n,
        total_prop = total_val$pct,
        total_sd = NA
      ))
    }
  }

  # Race/Ethnicity
  if (!is.null(table_data$race)) {
    for (i in 1:nrow(table_data$race)) {
      level <- table_data$race$level[i]
      lapse_val <- parse_n_pct(table_data$race$stat_lapse[i])
      no_lapse_val <- parse_n_pct(table_data$race$stat_no_lapse[i])
      total_val <- parse_n_pct(table_data$race$stat_total[i])

      export_df <- rbind(export_df, data.frame(
        variable = paste0("Race_", level),
        lapse_value = lapse_val$n,
        lapse_prop = lapse_val$pct,
        lapse_sd = NA,
        no_lapse_value = no_lapse_val$n,
        no_lapse_prop = no_lapse_val$pct,
        no_lapse_sd = NA,
        total_value = total_val$n,
        total_prop = total_val$pct,
        total_sd = NA
      ))
    }
  }

  # Insurance
  if (!is.null(table_data$insurance)) {
    for (i in 1:nrow(table_data$insurance)) {
      level <- table_data$insurance$level[i]
      lapse_val <- parse_n_pct(table_data$insurance$stat_lapse[i])
      no_lapse_val <- parse_n_pct(table_data$insurance$stat_no_lapse[i])
      total_val <- parse_n_pct(table_data$insurance$stat_total[i])

      export_df <- rbind(export_df, data.frame(
        variable = paste0("Insurance_", level),
        lapse_value = lapse_val$n,
        lapse_prop = lapse_val$pct,
        lapse_sd = NA,
        no_lapse_value = no_lapse_val$n,
        no_lapse_prop = no_lapse_val$pct,
        no_lapse_sd = NA,
        total_value = total_val$n,
        total_prop = total_val$pct,
        total_sd = NA
      ))
    }
  }

  # DR Severity
  if (!is.null(table_data$dr)) {
    for (i in 1:nrow(table_data$dr)) {
      level <- table_data$dr$level[i]
      lapse_val <- parse_n_pct(table_data$dr$stat_lapse[i])
      no_lapse_val <- parse_n_pct(table_data$dr$stat_no_lapse[i])
      total_val <- parse_n_pct(table_data$dr$stat_total[i])

      export_df <- rbind(export_df, data.frame(
        variable = paste0("DR_", level),
        lapse_value = lapse_val$n,
        lapse_prop = lapse_val$pct,
        lapse_sd = NA,
        no_lapse_value = no_lapse_val$n,
        no_lapse_prop = no_lapse_val$pct,
        no_lapse_sd = NA,
        total_value = total_val$n,
        total_prop = total_val$pct,
        total_sd = NA
      ))
    }
  }

  # Treatments
  treatment_names <- c("PRP", "Focal_laser", "Anti_VEGF", "Other_injections")
  treatment_data <- list(table_data$treatments$prp, table_data$treatments$focal,
                        table_data$treatments$vegf, table_data$treatments$other)

  for (j in 1:length(treatment_names)) {
    if (!is.null(treatment_data[[j]])) {
      tx_data <- treatment_data[[j]]
      export_df <- rbind(export_df, data.frame(
        variable = treatment_names[j],
        lapse_value = tx_data$n[tx_data$lapse_group == "Lapse in care"],
        lapse_prop = tx_data$pct[tx_data$lapse_group == "Lapse in care"],
        lapse_sd = NA,
        no_lapse_value = tx_data$n[tx_data$lapse_group == "No lapse in care"],
        no_lapse_prop = tx_data$pct[tx_data$lapse_group == "No lapse in care"],
        no_lapse_sd = NA,
        total_value = tx_data$n[tx_data$lapse_group == "Total"],
        total_prop = tx_data$pct[tx_data$lapse_group == "Total"],
        total_sd = NA
      ))
    }
  }

  # Comorbidities
  comorb_names <- c("Glaucoma", "Prior_cataract_surgery", "Other_retinal_disease")
  comorb_data <- list(table_data$comorbidities$glaucoma, table_data$comorbidities$cataract,
                     table_data$comorbidities$retina)

  for (j in 1:length(comorb_names)) {
    if (!is.null(comorb_data[[j]])) {
      com_data <- comorb_data[[j]]
      export_df <- rbind(export_df, data.frame(
        variable = comorb_names[j],
        lapse_value = com_data$n[com_data$lapse_group == "Lapse in care"],
        lapse_prop = com_data$pct[com_data$lapse_group == "Lapse in care"],
        lapse_sd = NA,
        no_lapse_value = com_data$n[com_data$lapse_group == "No lapse in care"],
        no_lapse_prop = com_data$pct[com_data$lapse_group == "No lapse in care"],
        no_lapse_sd = NA,
        total_value = com_data$n[com_data$lapse_group == "Total"],
        total_prop = com_data$pct[com_data$lapse_group == "Total"],
        total_sd = NA
      ))
    }
  }

  # Continuous variables
  cont_vars <- list(
    list(name = "Baseline_VA_logMAR", data = table_data$continuous$va),
    list(name = "DCSI", data = table_data$continuous$dcsi),
    list(name = "CCI", data = table_data$continuous$cci)
  )

  for (cv in cont_vars) {
    if (!is.null(cv$data)) {
      lapse_stats <- parse_mean_sd(cv$data$stat[cv$data$lapse_group == "Lapse in care"])
      no_lapse_stats <- parse_mean_sd(cv$data$stat[cv$data$lapse_group == "No lapse in care"])
      total_stats <- parse_mean_sd(cv$data$stat[cv$data$lapse_group == "Total"])

      export_df <- rbind(export_df, data.frame(
        variable = cv$name,
        lapse_value = lapse_stats$mean,
        lapse_prop = NA,
        lapse_sd = lapse_stats$sd,
        no_lapse_value = no_lapse_stats$mean,
        no_lapse_prop = NA,
        no_lapse_sd = no_lapse_stats$sd,
        total_value = total_stats$mean,
        total_prop = NA,
        total_sd = total_stats$sd
      ))
    }
  }
